---
title: "CDI_analysis"
output: html_document
---
#A breakdown of the CDM First Word Survey by CDI classification of the word. Some words were omitted due to not being translatable, and some of the words were classified in an ad-hoc manner

---
#Preliminaries 
```{r}
library(plyr)
library(ggplot2)
library(reshape2)
theme_set(theme_bw())
df = read.csv("CDM_CDIs.csv") 

```

---
#Freqs of First words - count
```{r}
freqs <- ddply(df, .(Word), summarise, count=length(Word))
freqs$Word <- factor(freqs$Word, levels=with(freqs, Word[order(count, Word, decreasing = TRUE)]))
q <- qplot(Word, count, geom="bar", stat="identity", 
      data=subset(freqs, count>5))
q + theme(axis.text.x = element_text(angle=90, hjust = 1))
```

#Freqs of first words props 
```{r}
freqs <- ddply(df, .(Word), summarise, count=length(Word))
freqs$Word <- factor(freqs$Word, levels=with(freqs, Word[order(count, Word, decreasing = TRUE)]))
freqs$prop <- freqs$count/sum(freqs$count)
q <- qplot(Word, prop, geom="bar", stat="identity", 
      data=subset(freqs, count>5))
q + theme(axis.text.x = element_text(angle=90, hjust = 1))
```
---
#Prop of first word by gender 
```{r}
freqs <- ddply(df, .(Word,gender), summarise, count=length(Word))
freqs <- ddply(freqs, .(Word), mutate, total.count = sum(count))
freqs$Word <- factor(freqs$Word, 
                     levels=unique(with(freqs, 
                                 Word[order(total.count, 
                                            Word, decreasing = TRUE)])))
freqs$prop <- freqs$count/sum(freqs$count)
q <- qplot(Word, prop, fill=gender, 
           position="dodge",
           geom="bar",
           stat="identity",
           data=subset(freqs, total.count>5 & gender!=""))
q + theme(axis.text.x = element_text(angle=90, hjust = 1))

```
---

#Freqs of CDIs - Count
```{r}
freqs <- ddply(df, .(CDI_category), summarise, count=length(CDI_category))
freqs$CDI_category <- factor(freqs$CDI_category, 
                     levels=with(freqs, 
                                 CDI_category[order(count, CDI_category, decreasing = TRUE)]))
q <- qplot(CDI_category, count, geom="bar", stat="identity", 
      data=subset(freqs))
q + theme(axis.text.x = element_text(angle=90, hjust = 1))

               
```
--- 
#Proportions of CDIS 
```{r}
freqs <- ddply(df, .(CDI_category), summarise, count=length(CDI_category))
freqs$CDI_category <- factor(freqs$CDI_category, 
                     levels=with(freqs, 
                                 CDI_category[order(count, CDI_category, decreasing = TRUE)]))
freqs$prop <- freqs$count/sum(freqs$count)

q <- qplot(CDI_category, prop, geom = "bar", stat="identity", data=subset(freqs, count>1))+theme_bw()
q + theme(axis.text.x = element_text(angle=90, hjust = 1))


```

--- 
#Proportions with Ages 
```{r}
freqs <- ddply(df, .(CDI_category, Age_utterance), summarise, count=length(CDI_category))
freqs <- ddply(freqs, .(CDI_category), mutate, total.count=sum(count))
freqs <- ddply(freqs, .(Age_utterance), mutate, prop = count/sum(count))
freqs$prop <- freqs$count/sum(freqs$count)

freqs$CDI_category <- factor(freqs$CDI_category, 
                             levels=unique(with(freqs, 
                                                CDI_category
                                                [order(total.count, CDI_category, decreasing = TRUE)])))

q <- qplot(CDI_category, prop, fill=Age_utterance, 
      position="dodge", geom="bar", stat="identity", data=subset(freqs, total.count>1 & 
                                                                   Age_utterance!="" &
                                                                   Age_utterance!="dontremember"))

q + theme(axis.text.x = element_text(angle=90, hjust = 1))

```
---
#Proportions by gender 
```{r}
freqs <- ddply(df, .(CDI_category, gender), summarise, count=length(CDI_category))
freqs <- ddply(freqs, .(CDI_category), mutate, total.count=sum(count))
freqs <- ddply(freqs, .(gender), mutate, prop = count/sum(count))
freqs$prop <- freqs$count/sum(freqs$count)

freqs$CDI_category <- factor(freqs$CDI_category, 
                             levels=unique(with(freqs, 
                                                CDI_category
                                                [order(total.count, CDI_category, decreasing = TRUE)])))

q <- qplot(CDI_category, prop, fill=gender, 
      position="dodge", geom="bar", stat="identity", data=subset(freqs, total.count>1 & gender!=""))
                                                                  
q + theme(axis.text.x = element_text(angle=90, hjust = 1))

```
--- 
#Age Split - 10m against everyone else 
```{r}
df$ten <- df$Age_utterance == "10mOrBefore"
#breaking up each age group individually, just in case we want to do something with it later 
df$eleven <- df$Age_utterance == "11m"
df$twelve <- df$Age_utterance == "12m"
df$thirteen <- df$Age_utterance == "13m"
df$fourteen <- df$Age_utterance == "14mOrLater"

freqs <- ddply(df, .(CDI_category,ten), summarise, count=length(CDI_category))
freqs <- ddply(freqs, .(CDI_category), mutate, total.count = sum(count))
freqs <- ddply(freqs, .(ten), mutate, prop = count / sum(count))

freqs$CDI_category <- factor(freqs$CDI_category,
                             levels=unique(with(freqs, 
                                                CDI_category[order(total.count, CDI_category, decreasing = TRUE)])))
freqs$prop <- freqs$count/sum(freqs$count)

q <- qplot(CDI_category, prop, fill=ten, 
      position="dodge",
      geom="bar", stat="identity", 
      data=subset(freqs,total.count>1))
q + theme(axis.text.x = element_text(angle=90, hjust = 1))
```

